Very simplified likelihood

### 2023-11-26 inst/very_simplified_likelihood.Rmd
consistentification with jss_hyper3

Consider a race between $n+1$ competitors of strength $a,b,\ldots,b$
where we require the total strength to be unity: $a+nb=1$.  Suppose
our bro comes first:

@@

Comes second:

\[
\mathcal{L}(a)=b\cdot\frac{a}{1-b}
\]

Third:

\[
\mathcal{L}(a)=b\cdot\frac{b}{1-b}\cdot\frac{a}{1-2b}
\]

Fourth:

\[
\mathcal{L}(a)=b\cdot\frac{b}{1-b}\cdot\frac{b}{1-2b}\cdot\frac{a}{1-3b}
\]

Placing after $r$ of the $b$ clones finish:

\[
\mathcal{L}(a)=\frac{b^ra}{(1-b)(1-2b)\cdots(1-rb)}=(1-nb)\frac{(-1)^r\Gamma(1-1/b)}{\Gamma(1+r-1/b)}
\]

@@

```{r solvescore}
  MLE <- function(r,n){
  small <- 1e-6
  a <- optimize(function(b){
       f_single_direct(b,r=r,n=n,log=TRUE)},
       maximum=TRUE, c(small,1-small))$maximum
  b <- (1-a)/n
  print(c(a,b))
  return(a/(a+b))
}

MLE(0,10)
MLE(1,10)
MLE(2,10)
MLE(3,10)
MLE(4,10)
MLE(5,10)
MLE(6,10)

MLE(0,11)
MLE(1,11)
MLE(2,11)
MLE(3,11)
MLE(4,11)
MLE(5,11)
MLE(6,11)

>>>>

Consider a race between a competitor of Bradley-Terry strength $a$ and
$n$ competitors each of strength $b$; we require $a+b=1$.  An
observation is indexed by $r$, the number of $b$'s finishing ahead of
$a$, so $0\leqslant r\leqslant n$.

### Likelihood functions

Suppose our bro comes first (zero $b$'s ahead of him):

@@

Comes second, one $b$ ahead of him:

\[
\mathcal{L}(a)=b\cdot\frac{a}{1-b}=b
\]

Third, two ahead of him:

\[
\mathcal{L}(a)=\frac{b}{1}\cdot\frac{b}{1-b}\cdot\frac{a}{1-2b}=\frac{b^2}{1-2b}
\]

Fourth, three ahead:

\[
\mathcal{L}(a)=b\cdot\frac{b}{1-b}\cdot\frac{b}{1-2b}\cdot\frac{a}{1-3b}=\frac{b^3}{(1-2b)(1-3b)}
\]

Fifth, four ahead:

\[
\mathcal{L}(a)=b\cdot\frac{b}{1-b}\cdot\frac{b}{1-2b}\cdot\frac{b}{1-3b}\cdot\frac{a}{1-4b}=\frac{b^4}{(1-2b)(1-3b)(1-4b)}
\]

Placing after $r$ of the $n$ clones of strength $b$ finish:

\[
\mathcal{L}(b)=\frac{b^ra}{(1-b)(1-2b)\cdots(1-rb)}=\frac{b^r}{(1-2b)\cdots(1-rb)},\qquad r\geqslant 2
\]

\[
\mathcal{L}(b)=\frac{b^{r+1}}{(1-2b)\cdots(1-rb)}=\frac{(-1)^{r+1}b\Gamma(2-1/b)}{\Gamma(1+r-1/b)}
\]

The case $n=2$ was considered by Hankin 2023.

@@

```{r solvescore}
  MLE <- function(r,n){
  stopifnot(length(r)==1)
  small <- 1e-6
  a <- optimize(function(b){
       f_single_direct(b,r=r,n=n,log=TRUE)},
       maximum=TRUE, c(small,1-small))$maximum
  b <- (1-a)/n
#  print(c(a,b))
  return(c(a/(a+b),b/(a+b)))
}
```

```{r rbindmle}
o <-
rbind(
r0=MLE(0,9),
r1=MLE(1,9),
r2=MLE(2,9),
r3=MLE(3,9),
r4=MLE(4,9),
r5=MLE(5,9),
r6=MLE(6,9),
r7=MLE(7,9),
r8=MLE(8,9),
r9=MLE(9,9))
o
matplot(o)
```

## Compare New Zealand with Australia

```{r nzaus}
x <- as.matrix(read.table("olympiad.txt",header=TRUE))
x <- x[c("AUS","NZL"),]
x
```
